\vspace{-1em}Guidance, please! Towards a framework for RDF-based constraint languages.

### 2015-07-16 paper/paper 2/dc-2015.tex
3.2

There are simple constraints which are not expressible in DL but can still be described using the RDF-CV such as constraints of the type \emph{literal pattern matching} (\emph{R-44}) which restrict literals to match given patterns. The \emph{universal quantification (R-91)} {\small\ms{Book $\sqsubseteq$ $\forall$ identifier.ISBN}} ensures that books can only have valid \emph{ISBN} identifiers, i.e., strings that match a given regular expression.
%Even though, the restriction of the datatype \emph{ISBN} cannot be expressed in DL, \emph{OWL 2 DL} can be used to express the \emph{literal pattern matching} constraint:
%\begin{ex}
%ISBN a RDFS:Datatype ; owl:equivalentClass [ a RDFS:Datatype ;
    %owl:onDatatype xsd:string ; 
    %owl:withRestrictions ([ xsd:pattern "^\d{9}[\d|X]" ])] .
%\end{ex} 
%The first OWL 2 axiom explicitly declares {\em ISBN} to be a datatype. %\er{Why start from second and then first? Also better we say "owl axiom" when we talk axiom of owl, to prevent ambiguity.}\tb{resolved}
%The second OWL 2 axiom defines {\em ISBN} as an abbreviation for a datatype restriction on {\em xsd:string}. 
%The datatype {\em ISBN} can be used just like any other datatype like in the \emph{universal restriction} above.
%There are multiple use cases associated with the requirement to match literals according to given patterns (\emph{Literal Pattern Matching}\footnote{Corresponds to \emph{R-44-PATTERN-MATCHING-ON-RDF-LITERALS}}).
%The enterprise vessel, e.g.,  can only have the registry numbers "NCC-1701", "NCC-1701-A", "NCC-1701-B", "NCC-1701-C", "NCC-1701-D", or "NCC-1701-E".
%The universal restriction can be represented in DL:
%\ms{Enterprise $\sqsubseteq$ $\forall$ registryNumber.RegistryNumber}.
%The restriction of the datatype \ms{RegistryNumber}, however, cannot be expressed in DL, but OWL 2 DL can be used anyway to express the literal pattern matching constraint:
%
%\begin{ex}
%RegistryNumber
    %a rdfs:Datatype ; owl:equivalentClass [ a rdfs:Datatype ;
        %owl:onDatatype xsd:string ;
        %owl:withRestrictions ( [ xsd:pattern "NCC-1701([-][A-E])?" ] ) ] .
%\end{ex}
%The second axiom defines \ms{RegistryNumber} as an abbreviation for a datatype restriction on \ms{xsd:string}. 
%The first axiom explicitly declares \ms{RegistryNumber} to be a datatype. 
%The datatype \ms{RegistryNumber} can be used just like any other datatype like in the universal restriction above.
%The \emph{literal pattern matching} constraint validates \emph{ISBN} literals according to the regular expression causing constraint violations for triples which do not not match. 
%\ms{Janeway commandsEnterprise Voyager} and \ms{Voyager registryNumber "NCC-74656"\textasciicircum{}\textasciicircum{}RegistryNumber}, 
%but not for the triples \ms{Picard commandsEnterprise Enterprise} and \ms{Enterprise registryNumber "NCC-1701-E"\textasciicircum{}\textasciicircum{}RegistryNumber}.

Even though, constraints of the type \emph{literal pattern matching} cannot be expressed in DL, OWL 2 can be used to formulate the constraint:

@@

Table \ref{tab:simple-constraint-not-expressible-in-dl)} presents (1) in the first line how the literal pattern matching simple constraint which is not expressible in DL and (2) in the second line how the universal quantification complex constraint which is expressible in DL are represented using the RDF-CV. Thereby, the context class \emph{ISBN}, whose instances must satisfy the simple constraint, is reused within the list of classes the complex constraint refers to. The literal pattern matching constraint type introduces the constraining element \emph{REGEX} whose validation has to be implemented once like for any other constraining element.
%In general, validation has to be implemented once for each \emph{generic constraint type} which is not expressible in DL. 
%\begin{gcotable}
%property & Enterprise & registryNumber & - & RegistryNumber & $\forall$ & - \\
%class & RegistryNumber & - & - & xsd:string & REGEX & 'NCC-1701([-][A-E])?' \\
%\end{gcotable}
\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Simple Constraints which are not Expressible in DL}
	\label{tab:simple-constraint-not-expressible-in-dl)}
	\centering
		\begin{tabular}{c|c|c|c|c|c}
      \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
      ISBN & - & - & xsd:string & REGEX & '\string^\text{$\backslash$d$\{9\}$[$\backslash$d$\mid$X]}\$' \\
      Book & identifier & - & ISBN & universal quantification & - \\
		\end{tabular}
\end{table} %\tb{Erman, can you please format the regular expression in the table above correctly?}

>>>>

Constraints of 36\% of the constraint types are not expressible in DL but can still be described using the RDF-CV such as constraints of the type \emph{literal pattern matching} (\emph{R-44}) which restrict literals to match given patterns. The \emph{universal quantification (R-91)} {\small\ms{Book $\sqsubseteq$ $\forall$ identifier.ISBN}} ensures that books can only have valid \emph{ISBN} identifiers, i.e., strings that match a given regular expression.
%Even though, the restriction of the datatype \emph{ISBN} cannot be expressed in DL, \emph{OWL 2 DL} can be used to express the \emph{literal pattern matching} constraint:
%\begin{ex}
%ISBN a RDFS:Datatype ; owl:equivalentClass [ a RDFS:Datatype ;
    %owl:onDatatype xsd:string ; 
    %owl:withRestrictions ([ xsd:pattern "^\d{9}[\d|X]" ])] .
%\end{ex} 
%The first OWL 2 axiom explicitly declares {\em ISBN} to be a datatype. %\er{Why start from second and then first? Also better we say "owl axiom" when we talk axiom of owl, to prevent ambiguity.}\tb{resolved}
%The second OWL 2 axiom defines {\em ISBN} as an abbreviation for a datatype restriction on {\em xsd:string}. 
%The datatype {\em ISBN} can be used just like any other datatype like in the \emph{universal restriction} above.
%There are multiple use cases associated with the requirement to match literals according to given patterns (\emph{Literal Pattern Matching}\footnote{Corresponds to \emph{R-44-PATTERN-MATCHING-ON-RDF-LITERALS}}).
%The enterprise vessel, e.g.,  can only have the registry numbers "NCC-1701", "NCC-1701-A", "NCC-1701-B", "NCC-1701-C", "NCC-1701-D", or "NCC-1701-E".
%The universal restriction can be represented in DL:
%\ms{Enterprise $\sqsubseteq$ $\forall$ registryNumber.RegistryNumber}.
%The restriction of the datatype \ms{RegistryNumber}, however, cannot be expressed in DL, but OWL 2 DL can be used anyway to express the literal pattern matching constraint:
%
%\begin{ex}
%RegistryNumber
    %a rdfs:Datatype ; owl:equivalentClass [ a rdfs:Datatype ;
        %owl:onDatatype xsd:string ;
        %owl:withRestrictions ( [ xsd:pattern "NCC-1701([-][A-E])?" ] ) ] .
%\end{ex}
%The second axiom defines \ms{RegistryNumber} as an abbreviation for a datatype restriction on \ms{xsd:string}. 
%The first axiom explicitly declares \ms{RegistryNumber} to be a datatype. 
%The datatype \ms{RegistryNumber} can be used just like any other datatype like in the universal restriction above.
%The \emph{literal pattern matching} constraint validates \emph{ISBN} literals according to the regular expression causing constraint violations for triples which do not not match. 
%\ms{Janeway commandsEnterprise Voyager} and \ms{Voyager registryNumber "NCC-74656"\textasciicircum{}\textasciicircum{}RegistryNumber}, 
%but not for the triples \ms{Picard commandsEnterprise Enterprise} and \ms{Enterprise registryNumber "NCC-1701-E"\textasciicircum{}\textasciicircum{}RegistryNumber}.

Even though, constraints of the type \emph{literal pattern matching} cannot be expressed in DL, OWL 2 can be used to formulate this constraint:

@@

Table \ref{tab:simple-constraint-not-expressible-in-dl)} presents (1) how the not in DL expressible literal pattern matching constraint and (2) how the in DL expressible universal quantification are both represented using the RDF-CV. Thereby, the context class \emph{ISBN}, whose instances must satisfy the literal pattern matching constraint, is reused within the list of classes the universal quantification refers to. The literal pattern matching constraint type introduces the constraining element \emph{REGEX} whose validation has to be implemented once like for any other constraining element.
%In general, validation has to be implemented once for each \emph{generic constraint type} which is not expressible in DL. 
%\begin{gcotable}
%property & Enterprise & registryNumber & - & RegistryNumber & $\forall$ & - \\
%class & RegistryNumber & - & - & xsd:string & REGEX & 'NCC-1701([-][A-E])?' \\
%\end{gcotable}
\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Simple Constraints which are not Expressible in DL}
	\label{tab:simple-constraint-not-expressible-in-dl)}
	\centering
		\begin{tabular}{c|c|c|c|c|c}
      \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
      ISBN & - & - & xsd:string & REGEX & '\string^\text{$\backslash$d$\{9\}$[$\backslash$d$\mid$X]}\$' \\
      Book & identifier & - & ISBN & universal quantification & - \\
		\end{tabular}
\end{table} %\tb{Erman, can you please format the regular expression in the table above correctly?}
